---
title: "Day 10: Morning tasks & activities - KEY"
subtitle: "Fiddler crabs and Bergmann’s rule"
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
library(tidyverse)
library(skimr)
library(kableExtra)
library(lterdatasampler)
```

## GitHub collaboration with Yellow fiddler crab data

1. Read `fiddler_crab.csv` using `read_csv()`

::: {.callout-important}
## Solution
```{r}
#| include: false
fiddler_crab <- read_csv(here::here("course-materials", "eod-practice", "data", "fiddler_crab.csv"))
```

```{r}
#| eval: false
fiddler_crab <- read_csv(here::here("data", "fiddler_crab.csv"))
```
:::

2. Create an exploratory (unfinalized) plot of fiddler crab carapace widths observed at the different latitudes

::: {.callout-important}
## Solution
```{r}
ggplot(data = fiddler_crab) +
  geom_point(aes(y = Latitude, x = carapace_width))
```
:::

3. Add a figure caption using `#| fig-cap: "this is my caption"` in the code chunk where the graph is created. Update code chunk options (hint: `execute: ` in YAML) so that only your finalized graph and figure caption show up in your knitted report (i.e., no code should show up)

::: {.callout-important}
## Solution
```{r}
#| fig-cap: "Fiddler crab carapace widths observed at the different latitudes"
ggplot(data = fiddler_crab) +
  geom_point(aes(y = Latitude, x = carapace_width))
```
:::

4. Add an unfinalized summary table containing the mean, standard deviation, and sample size of fiddler crab carapace widths by site (tip: use `round(mean(), 2)` to round a value to 2 decimal places)

::: {.callout-important}
## Solution
```{r}
fiddler_summary <- fiddler_crab |> 
  drop_na(carapace_width) |> 
  group_by(Site) |> 
  summarize(mean_carapace = round(mean(carapace_width), 2),
            sd_carapace = round(sd(carapace_width), 2),
            n = n()) |> 
  ungroup()
```
:::

5. **Finalize the summary table** a bit so that it looks more polished when rendered, including with updated column names (including units as relevant)

::: {.callout-important}
## Solution
```{r}
fiddler_summary |> 
  kable(col.names = c("Site code", "Mean carapace width (mm)", "Standard deviation (mm)", "Sample size")) |> 
  kable_styling(full_width = FALSE)
```
:::
