---
title: "Day 10b: Afternoon tasks & activities - KEY"
subtitle: "Long-term ice cover duration: North Temperate Lakes LTER"
execute:
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(lubridate)
library(here)
library(kableExtra)
```

## Data wrangling and visualization of 

1. Read in both files (ice cover duration and meteorological data). Note that the ice cover data has some NA values stored as `-999`

::: {.callout-important}
## Solution
```{r}
#| include: false
icecover <- read_csv(here("course-materials", "eod-practice", "data", "ntl33_v7.csv"), na = "-999")
weather <- read_csv(here("course-materials", "eod-practice", "data", "ntl20_v6.csv"))
```

```{r}
#| eval: false
icecover <- read_csv(here("data", "ntl33_v7.csv"), na = "-999")
weather <- read_csv(here("data", "ntl20_v6.csv"))
```
:::

2. Create a visualization that compares ice duration across different lakes in the Madison Lake Area. Include both the actual ice duration values *and* some representation of summary statistics for each lake.

::: {.callout-important}
## Solution
```{r}
ggplot(data = icecover, aes(x = lakeid, y = ice_duration)) +
  geom_jitter(width = 0.2, color = "gray50") +
  geom_boxplot(width = 0.3, fill = NA, size = 1, outlier.color = NA)
```
:::

3. Find the mean ice cover duration in Madison Lake Area by year (i.e., you'll only be grouping by the year variable here). Understanding check: what are you finding the mean of here? Why are there multiple observations for each year?)

::: {.callout-important}
## Solution
```{r}
mean_duration <- icecover  |>
  group_by(year4) |> 
  summarize(mean_duration = mean(ice_duration, na.rm = TRUE)) |> 
  ungroup()
```
:::

4. Create an exploratory visualization of mean ice cover duration by year for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe. 
  
::: {.callout-important}
## Solution
```{r}
ggplot(data = mean_duration, aes(x = year4, y = mean_duration)) +
  geom_line()
```
:::

5. Find the mean air temperature (using the daily average air temperature - adjusted) in Madison Lake Area **by year** using **only observations from winter months (December, January, February)**
  
::: {.callout-important}
## Solution
```{r}
weather_summary <- weather |> 
  filter(month %in% c(12, 1, 2)) |> 
  group_by(year4) |> 
  summarize(mean_temp = mean(ave_air_temp_adjusted, na.rm = TRUE)) |> 
  ungroup()
```
:::

6. Create an exploratory visualization of mean winter temperatures for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.
  
::: {.callout-important}
## Solution
```{r}
ggplot(data = weather_summary, aes(x = year4, y = mean_temp)) +
  geom_line()
```
:::

7. Join the mean winter air temperatures to the mean ice cover duration data you found above

::: {.callout-important}
## Solution
```{r}
ice_weather_join <- mean_duration |> 
  full_join(weather_summary)
```
:::

8. Create an exploratory scatterplot of mean winter temperatures versus mean ice duration. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe. 
  
::: {.callout-important}
## Solution
```{r}
ggplot(data = ice_weather_join, aes(x = mean_temp, y = mean_duration)) +
  geom_point()
```
:::

9. Explore thaw dates since 1970. Create a subset of the ice cover duration data since 1970 (ok to include 1970), then convert the `ice_off` column to a `Date`. Use `lubridate::yday()` to pull the numeric *day of the year* from that Date. Create an exploratory plot of the ice off day (numeric day-of-the-year) from 1970 - present. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe. 

::: {.callout-important}
## Solution
```{r}
thaw_dates <- icecover |> 
  filter(year4 >= 1970) |> 
  mutate(ice_off = lubridate::mdy(ice_off)) |> 
  mutate(day_ice_off = lubridate::yday(ice_off))

ggplot(data = thaw_dates, aes(x = year4, y = day_ice_off)) +
  geom_point()
```
:::
